The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-04-21 Paper/jaccardbaetica.tex
comentarios

%primer párrafo más general---- especifico (¿dónde se ha aplicado?)
Material culture is one of the most frequent indicators of trade in the archaeological record. In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}. Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}. Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places. Roman provinces such as \textit{Baetica} (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire. Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}. As a consequence, this product was distributed in a large number of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.

@@

%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres. To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}. Stamps will be used to identify economic patterns by analysing their similarity. If workshops and provinces share stamps with similar traits, then we can identify connections. By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.

@@

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.

Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.

@@

@@

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}). However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

%primer párrafo más general---- especifico (¿dónde se ha aplicado?)
\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}

Material culture is one of the most frequent indicators of trade in the archaeological record. In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}. Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico noÇ?}. Given this fact, an important mechanism of production control under the Roman government was created spreading by different areas attending to the richness of different places. Roman provinces such as \textit{Baetica} (currently Andalusia, south Spain) became important production centres of olive oil during the Roman Empire. Olive oil was considered the liquid gold since it was used in different aspect of daily life for cooking, hygiene or lighting  \citep{mattingly_d.j._oil_1988}. As a consequence, this product was distributed in a large number of amphorae along the province, mostly to supply the Roman Army and Italy \citep{blazquez_exportacion_1980}.

@@

%This paper explore--- resumir en dos líneas el objeto de estudio y definir que ha sido usado para ello dispersion of culture
\xavi{este aim es demasiado general}
This paper aims to understand the production dynamics in connection with different areas within the Roman Empire. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres. To do this, an Ecology approach has been used to analyse the dispersion of stamps between amphora workshops \citep{rubio-campillo_ecology_2018}\xavi{lo de Ecology approach es raro no? Yo lo usé pero en el título...quizás population approach?}. Stamps will be used to identify economic patterns by analysing their similarity. If workshops and provinces share stamps with similar traits, then we can identify connections. By contrast, if we do not detect similar stamps between workshops, then workshops worked independently.

\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}

@@

\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}

\xavi{seguro que tiene subsecciones este journal?}

@@

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be affected by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamp in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province. \xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}

Therefore, this study proposes an evolutionary baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance. 
\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}

@@

\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}

@@

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC database \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}) \xavi{no fue la database la que compiló los sellos, no? :-D}}. However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}

@@

\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}

@@

\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro. Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}

@@

\xavi{pos eso, los mapas...te parece si los hago yo...? andaaa vaaaa déjame hacer algo}

@@

\xavi{yo creo que solo hay 1 nivel en el JAMT, ni sub ni subsub}

@@

\xavi{pero este es el único método que usas?}

@@

\xavi{no puede ser vertical la figura? se leería mejor}

@@

\xavi{no tengo claro que le puedas decir correlación, porque eso es un valor específico entre 2 distribuciones numéricas. Te refieres a similaridad?}
